---
title: "Personal Consumption Structure in each State"
author: "You Shan Shen"
output: 
  flexdashboard::flex_dashboard:
    orientation: columns
    vertical_layout: fill
---

```{r setup,, include = FALSE, echo = FALSE}
library(tidyverse)
library(lubridate)
library(gt)
library(paletteer)
library(ggridges)
library(plotly)
library(sf)
library(ggmap)
library(tidycensus)
library(leaflet)
library(osmdata)
library(tigris)
library(ggthemes)
library(formattable)
library(leaflegend)
library(knitr)
knitr::opts_knit$set(echo = TRUE, 
root.dir = rprojroot::find_rstudio_root_file())
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, error = FALSE, message = FALSE)

knitr::opts_chunk$set(
  fig.width = 6,
  out.width = "90%",
 fig.align  = "center",
  cache = FALSE,
  eval  = TRUE,
  echo  = TRUE,
  warning = FALSE
)


theme_set(theme(plot.title = element_text(hjust = 0.5), legend.position = "right"))
```

```{r read files, include = FALSE}
file_dir  <- "./data/SAGDP"
data_files <- list.files(file_dir)
file_paths <- str_c(file_dir, data_files) 

t1_file_list <- list.files(file_dir, pattern = "SAGDP2N")
t1_file_list <- t1_file_list[!str_detect(t1_file_list, c("AREAS", "definition", "Footnotes" , "modify"))]

personal_consumption.df <- tibble(
           file_dir   = "./data/SAGDP/", 
           file_list  = t1_file_list, 
           state      = map(str_split(string = file_list, "_"), 2),
           file_paths = str_c(file_dir, file_list)) %>%
  dplyr::select(state, file_paths)

personal_consumption_data <- personal_consumption.df %>%
  mutate( 
    data = map(.x = file_paths, ~read_csv(.x, show_col_types = FALSE, na = c("(L)","(D)","(NA)")))) %>%
  dplyr::select(-file_paths) %>%
  unnest() %>%
  filter(!LineCode %in% NA,
         !GeoName %in% 'United States *',
         !Description %in% c("All industry total","Private industries")) %>%
  select(!c(`1997`:`2009`) & !c(`2021`) & !c(`Unit`))

```

```{r, results = 'hide', fig.show='hide', include = FALSE}
library(DataExplorer)
data_explore_df  <- personal_consumption_data  
DataExplorer::introduce(data_explore_df)
DataExplorer::plot_missing(data_explore_df )
DataExplorer::plot_bar(data_explore_df )
DataExplorer::plot_histogram(data_explore_df)
```

```{r map legend color, include = FALSE }
 total_consumption <- personal_consumption_data %>% 
  nest(c(`2010`:`2018`)) %>%
  mutate(
    description_sum = map(.x = data, ~ .x %>% rowSums(na.rm = TRUE))) %>%
  unnest() %>%
  group_by(GeoName) %>%
  mutate( state_sum = sum(description_sum, na.rm = TRUE)) %>%
   select(c(GeoName,state_sum)) %>%
   unique() %>%
   arrange(desc(state_sum)) %>%
   filter(!(GeoName %in% c("Southeast","Far West","Mideast","Great Lakes","Southeast","Plains","Rocky Mountain","New England","Puerto Rico", "Southwest", "District of Columbia"))) %>%
   mutate(state_sum = state_sum/1000000) %>%
   mutate(
   GeoName = fct_reorder(GeoName, state_sum, .desc = FALSE)
  )

```
```{r before covid, include = FALSE}

pre_covid_percent_consumption_df <- personal_consumption_data %>% 
  nest(c(`2010`:`2018`)) %>%
  mutate(
    description_sum = map(.x = data, ~ .x %>% rowSums(na.rm = TRUE))) %>%
unnest() %>%
  group_by(GeoName) %>%
  mutate( state_sum = sum(description_sum, na.rm = TRUE)) %>%
   mutate( consumption_percent =  description_sum/state_sum,
          consumption_percent = consumption_percent*100) %>%
  mutate( across(c(consumption_percent ), ~ percent(.)/100)) %>%
  ungroup()

  
top3_comsumption <- pre_covid_percent_consumption_df  %>%
  group_by(GeoName) %>%
  arrange(desc(consumption_percent)) %>%
  group_by(GeoName) %>%
  slice(1:3)


review_precovid <- top3_comsumption %>%
  select( c(GeoName, Description, consumption_percent)) %>%
  rename( "Pre-COVID consumption %" = consumption_percent ) %>%
  mutate(
Description = case_when(
Description %in% "Private services-providing industries 3/" ~ "Providing industries",
Description %in% "Private goods-producing industries 2/" ~ "Producing industries",
Description %in% "Manufacturing and information" ~ "Manufacturing",
Description %in% "Government and government enterprises" ~ "Gov. enterprises",
Description %in% "Finance, insurance, real estate, rental, and leasing" ~ "Finance",
Description %in% "Federal civilian" ~ "Federal civilian",
Description %in% "Real estate and rental and leasing" ~ "Real estate",
Description %in% "Professional and business services" ~ "Business services",
Description %in% "Natural resources and mining" ~ "Natural resources",
Description %in% "Arts, entertainment, recreation, accommodation, and food services"~ "Arts, recreation", 
TRUE ~ Description
)) %>%
   mutate("Description" = str_c(`Description`, `Pre-COVID consumption %`, sep = " : ",collapse = NULL))

review_precovid %>% gt_preview()

```


```{r after covid, include = FALSE}

after_covid_percent_personal_consumption_df <- personal_consumption_data %>% 
  nest(c(`2019`:`2020`)) %>%
  mutate(
    description_sum = map(.x = data, ~ .x %>% rowSums(na.rm = TRUE))) %>%
unnest() %>%
  group_by(GeoName) %>%
  mutate( state_sum = sum(description_sum, na.rm = TRUE)) %>%
  mutate( consumption_percent =  description_sum/state_sum,
          consumption_percent = consumption_percent*100) %>%
   mutate( across(c(consumption_percent ), ~ percent(.)/100)) %>%
  ungroup()
  
  
after_covid_top3_comsumption <- after_covid_percent_personal_consumption_df %>%
  group_by(GeoName) %>%
  arrange(desc(consumption_percent)) %>%
  group_by(GeoName) %>%
  slice(1:3)

review_after_covid <- after_covid_top3_comsumption %>%
  select( c(GeoName, Description, consumption_percent)) %>%
    rename( "After COVID consumption %" = consumption_percent ) %>%
  mutate(
Description = case_when(
Description %in% "Private services-providing industries 3/" ~ "Providing industries",
Description %in% "Private goods-producing industries 2/" ~ "Producing industries",
Description %in% "Manufacturing and information" ~ "Manufacturing",
Description %in% "Government and government enterprises" ~ "Gov. enterprises",
Description %in% "Finance, insurance, real estate, rental, and leasing" ~ "Finance",
Description %in% "Federal civilian" ~ "Federal civilian",
Description %in% "Real estate and rental and leasing" ~ "Real estate",
Description %in% "Professional and business services" ~ "Business services",
Description %in% "Natural resources and mining" ~ "Natural resources",
TRUE ~ Description
)) %>%
   mutate("Description" = str_c(`Description`, `After COVID consumption %`, sep = " : ",collapse = NULL))

review_after_covid %>% gt_preview()
```





```{r leaflet setting, , include = FALSE}

us_state.df <- st_read("./data/state sf/cb_2020_us_state_5m.shp") %>%
  janitor::clean_names() %>%
  select(c("name", "geometry")) %>%
   filter(!(name %in% c("Puerto Rico"))) 


review_after_covid  %>% nest(-c(GeoName)) %>% distinct(GeoName)
review_precovid  %>% nest(-c(GeoName)) %>% distinct(GeoName)


covid_summary_top3 <- bind_cols(review_after_covid, review_precovid) %>%
  select(-c("GeoName...4"))  %>%
  rename( "After-COVID" = Description...2,
          "Pre-COVID" = Description...5 ) %>%
  mutate("Consumption Type" = str_c(`After-COVID`, `Pre-COVID`, sep = " | ",collapse = NULL),
         "Percent %" = str_c(`After COVID consumption %`, `Pre-COVID consumption %`, sep = " | ",collapse = NULL)) %>%
  select(c("GeoName...1", "Consumption Type", "Percent %")) %>%
  rename("State" = "GeoName...1") %>%
  select(-`Percent %`) %>%
  group_by(State)

leaflet_label.df <- reshape(
  transform( covid_summary_top3,
             q = ave(State, State, FUN = seq_along)),
  direction = "wide",
  idvar = "State",
  timevar = "q" )  %>%
  rename("1st Consumption" = "Consumption.Type.1",
         "2nd Consumption" = "Consumption.Type.2",
         "3rd Consumption" = "Consumption.Type.3")
  
              
us_geo.df <- left_join( us_state.df,
                        leaflet_label.df,
                         by = c( "name" = "State") ) %>%
  mutate( period = "After-Covid vs Pre-Covid")
  
  

library(viridis, quietly = TRUE)

us_label.df <- right_join(us_geo.df, total_consumption, by = c( "name" = "GeoName") ) %>%
  mutate( text_label = str_c("<br/>State: ", name,
                             "<br/>Period: ", period,
                             "<br/>1st Consumption :",`1st Consumption`,
                             "<br/>2nd Consumption :",`2nd Consumption`,
                             "<br/>3rd Consumption :",`3rd Consumption`)) %>%
  na.omit()
  

qpal <- colorNumeric(palette = "Oranges", domain = us_label.df$state_sum)
  
```
```{r leaflet, echo = FALSE, fig.width = 4, fig.height = 6}
leaflet(data = us_label.df ) %>%
  addTiles() %>%
  fitBounds(~-200,~18, ~-30,~80) %>%
  addProviderTiles('Esri.WorldStreetMap') %>%
  addPolygons(label = ~map(text_label, html),
              labelOptions = labelOptions(
                style = list("font-weight" = "normal", padding = "3px 8px"),
                textsize = "10px"),
              fillColor = ~qpal(us_label.df$state_sum),
              fillOpacity = 1,
              weight = 2,
              color = "IndianRed",
              dashArray = "3",
              highlightOptions = highlightOptions(
                weight = 3, 
                color = "black",
                dashArray = "", 
                fillOpacity = 1)) %>%
  addLegend(pal = qpal, values = ~us_label.df$state_sum, opacity = 0.7, title = "10yrs Personal Consumption Accumulation", position = "bottomright",
      labFormat = labelFormat(prefix = '', suffix = 'M', digits = 6, big.mark = ',', transform = identity)) 
````

***

PCE(Personal Consumption Expenditures) by state highlights

- Providing industries is the highest percentage in each state. It indicates that service industries are the most demanding aspect in all the consumption categories.
- Overall, compared with Pre-COVID, the top3 industry and their percentage didn’t change a lot in each state regardless of the total consumption.
- In recent ten years, the highest personal consumption expenditures states are California, New York, and Texas.



